An Information Theoretic Formalism \\for Intent Disambiguation

### 2019-02-22 TexStuff/IROS2018_ResubmitTex/GopinathArgall_IROS2019_2.tex
Version final - 2

\begin{abstract}
The effectiveness of assistive robots is closely related to their ability to infer the user's needs and intentions \textit{unambiguously} and provide appropriate assistance quickly and accurately. In this paper, we proposed a mode selection paradigm that enhances the autonomy's intent inference capabilities by exploiting the fact that how the human operates the robot and reveals his/her intent is intrinsically conditioned on the control mode in operation. We formulate this as a problem of intent disambiguation in information theoretic terms.  We propose two different methods for enhancing intent inference via improved disambiguation using the information theoretic concepts of \textit{entropy} and \textit{KL divergence}. In our system, the autonomy maintain a probability distribution over goals (beliefs) and we characterize the disambiguation capabilities of the different control modes/dimensions utilizing a model-based approach to compute the average information gain and content. 
%The proposed algorithms characterize the disambiguation capabilities of the \added{different control modes?} system through the forward projection of probability distributions over goals, utilizing an intent inference scheme and a kinematics model, and computing the average information gain and content. 
Previous work has shown that the success of a disambiguation algorithm depends on a variety of factors and parameters. To thoroughly investigate the impact of these various components, we present results from an extensive simulation-based study for both a point robot and physics-based simulation of a six degrees-of-freedom (DoF) robotic arm. Our results indicate that, compared to a baseline, the proposed disambiguation algorithms do enhance intent inference via faster intent disambiguation, which in turn allows autonomy assistance to step in earlier during task execution. We also find that goal inference is more accurate and the total amount of time assistance is engaged is higher.
\end{abstract}

@@

\section{Introduction}
Assistive machines such as robotic arms and smart wheelchairs have the potential to transform the lives of millions of people with motor impairments.~\citep{laplante1992assistive}. These machines can promote independence, enhance the quality of lives and revolutionize the way people interact with society. They can also help extend the mobility and manipulation capabilities of individuals, thereby helping motor-impaired people perform activities of daily living in a more effective manner.

An assistive robotic machine is typically controlled using an interface such as a joystick, a switch-based head array or a sip-and-puff that. These interfaces are low-dimensional, low-bandwidth and occasionally discrete and typically motion commands. For this reason, at any given point in time during task execution they can only operate in a subset of the entire control space. These subsets are referred to as \textit{control modes}~\citep{simpson2008tooth}. To schedule and execute switches between control modes can be both mentally and physically demanding, and might be alleviated in part by the introduction of robotics autonomy. The efficacy of such autonomy-endowed machines relies on their ability to infer the users' needs and intentions, and is often a bottleneck for providing appropriate assistance accurately, confidently and quickly. Due to the dimensionality mismatch between high-dimensional robots and low-dimensional interfaces, the user is constrained to produce control commands that likely do not express their true underlying intent unambiguously. That is, the control interfaces act like filters that restrict the amount of information regarding the user's intent that gets passed through to the machine and autonomy. 
%Furthermore, sparsity and noise in the control commands make the inference process even harder prompting the need for robust intent inference formalisms.

In the context of assistive robotic manipulation, the first step of a task is often to \textit{reach} towards discrete objects in the environment. Therefore in this domain, \textit{intent inference} can be cast as a problem of maintaining and updating the probability distributions over all possible discrete goals (objects) in the workspace upon receiving fresh evidence from the human actions and environment. In Bayesian terminology, maintaining and updating the belief is the \textit{recursive Bayesian filtering} problem. Intent inference algorithms typically rely on various environmental cues and task-relevant features such as the robot and goal positions, human control commands and biometric measures~\citep{croft2003estimating} as their evidence variables. In contrast, due to reasons of user adoption and cost our system infers intent exclusively based on the information contained in the constrained control commands issued to the assistive machine.

Our system is inspired by the following insights:
\begin{enumerate}
	\item At any given time, the user is constrained to operate the robot in a specific control mode. That is, the policy followed by the human while executing the reaching task is conditioned on the current active mode. 
	\item The posterior computation f the belief over goals depends on the human policy (treated as an evidence variable) which indirectly depends on the active control mode. 
	\item Utilizing a model-based approach to forward project the beliefs over goals, the autonomy can \textit{counterfactually } reason about the information gain and content for different control modes. 
	\item Using this information theoretic characterization of control modes the autonomy chooses the control mode with the highest intent disambiguation capability \textit{for} the human.
	\item By having the human provide control input within these information-rich control modes can likely improve the accuracy of intent inference. 
\end{enumerate}
%Our insight is that, at any given time, there is a specific set of control modes that are more informative \textit{for} the autonomy than others. Therefore, by having the human provide control input within these information-rich modes, the possibility of inferring the human's intent accurately and unambiguously is increased. \added{Talk about this specifically in terms of how mode selection can improve intent inference. }
This, in turn, allows the autonomy to step in and \textit{assist the human achieve their desired goal earlier}.  This is important, because when the intent inference mechanism infers the incorrect goal (or is simply too slow), the user and autonomy often end up competing instead of collaborating and can decrease user satisfaction and result in significantly worse performance. 
%than either the human or autonomy would produce on their own. 
In this paper, we address the problem of accurate intent inference by characterizing control modes according to the information content and proposing a mode switch assistance paradigm that selects the control mode in which a user-initiated motion will \textit{maximally disambiguate} their intent. We formalize the problem of intent disambiguation within the framework of information theory. Specifically, we make use of the information theoretic notions of entropy and Kullback-Leibler (KL) divergence to characterize and quantify the information content---with respect to intent disambiguation---of each control mode.

@@

Information theoretic approaches are widely utilized in the field of machine learning and robotics for optimal experiment design, for efficient data collection processes and for informing search strategies. Robot assistance schemes that seek to elicit more informative cues \textit{from} the human to clarify the underlying intent can be thought of as an information acquisition problem. Intent acquisition can leverage the underlying synergies and shared intentionality~\citep{tomasello2007shared} of human-robot teams and can be an active process in which the robot performs actions (for example, selecting a control mode or executing a robot motion) that will nudge the human to reveal her/his intent more clearly~\cite{sadigh2016information, sadigh2016planning}. Information theoretic metrics such as KL divergence can be utilized to identify regions of the sample space that will maximize information gain~\citep{tong2001active} and subsequently guide the data sampling process. Sensing robots designed for automated exploration and data acquisition tasks can benefit from exploring more information-rich regions in the environment~\citep{atanasov2014information}. If the spatial distribution of information density is known \textit{a priori}, information maximization can be accomplished by maximizing the ergodicity of the robot's trajectory with respect to the underlying information density map~\citep{miller2013trajectory, miller2016ergodic}.

For an assistive machine to provide appropriate kinds of assistance accurately and at the right time, it needs to have a good estimate of the human's underlying intent. Intent inference algorithms, therefore, play a vital role in the success of an assistive system. Intent inference and recognition can be classified into two broad categories: heuristic approaches and model-based approaches. Heuristic approaches are often simpler and computationally light-weight and seek to find direct mappings between various task relevant features (such as motion cues) and the human's underlying intention~\citep{baker2007goal, baker2009action}. On the other hand, in model-based approaches the system maintains a model of how a human maps states to control actions. The model can either be learned from data or can be hand-designed based on domain knowledge. For example, the human can be modeled within the Partially Observable Markov Decision Process (POMDP)~\citep{dragan2013policy, taha2011pomdp} framework and is assumed to behave according to a control policy that maps states to actions. However, in model-based approaches incorporating the entire history of states requires estimation of joint probability distributions over past states which can become computationally expensive and intractable quickly.

When there is a mismatch between the dimensionality of the problem and the control interface users have to continuously shift their focus from the task at hand to the choice of control mode during task execution, thereby resulting in a higher cognitive load. To solve this problem, various mode switching assistance paradigms have been proposed to alleviate task effort. For example, an automatic time-optimal mode switch assistance has been proposed which has shown to significantly improve user satisfaction~\citep{herlant2016assistive}. In the area of myoelectric prosthetics, dynamic switching approaches that learn the most effective control option during task execution using temporal difference and reinforcement learning have also been proposed~\cite{pilarski2012dynamic}.

@@

% Section~\ref{ssec:notation} outlines the mathematical notation used in this paper. Section~\ref{ssec:set_modes} presents a set-theoretic treatment of control modes and Section~\ref{ssec:disamb} describes the disambiguation algorithm. The mathematical details of the intent inference paradigms are outlined in detail in Section~\ref{ssec:inference}.
\subsection{Probability Distribution Over Goals}\label{ssec:notation}
 In assistive robotic manipulation, intent inference is most commonly the process of estimating the user's intended goal out of a set of discrete objects in the environment~\citep{calli2015ycb}. The set of all candidate goals is denoted by $\mathcal{G}$ with $n_g = \vert\mathcal{G}\vert$ and let $g^i$ refer to the $i^{th}$ goal with $i \in [1,2,\dots, n_g]$. 
% At any time $t$, the likelihood of goal $g$ being the desired goal follows a categorical probability distribution whose sample space is $\mathcal{G}$. 
% That is, 
% \begin{equation*}
% g \sim \textbf{Cat}(n_g | \boldsymbol{p}(t)) = \prod_{i=1}^{n_g}p^{i}(t)^{[g = g^i]}
% \end{equation*}
%where, $[g = g^i]$ is the Iverson bracket and evaluates to 1 if $g = g^i$, 0 otherwise. 
$\boldsymbol{p}(t)$ denotes the probability distribution over goals such that $\boldsymbol{p}(t) = [p_{g_1}^t, p_{g_2}^t,\dots, p_{g_{n_g}}^t)]^{T}$ where $p_{g_i}^t$ denotes the probability associated with goal $g_i$ at time $t$.  The probability $p_{g_i}^t$  also represents the robot's \textit{confidence} that goal $g_i$ is the human's intended goal at time $t$.

@@

 Let $\mathcal{M}$ denote the set of all control modes with $n_m = \vert\mathcal{M}\vert$. Additionally, let $m^i$ refer to the $i^{th}$ control mode where $i \in [1,2,\dots,n_m]$. Each control mode $m^i$ is a subset of $\mathcal{K}$ such that $\bigcup\limits_{i=1}^{n_m} m^i = \mathcal{K}$. The cardinality of each mode $m \in \mathcal{M}$, denoted by $\vert m \vert$ indicates the number of dimensions that can be controlled when operating in $m$.\footnote{Note that a dimension $k \in \mathcal{K}$ can be an element of multiple control modes.} Furthermore, the user can only operate in one of the $n_m$ control modes at any given time $t$. That is, for each $m \in \mathcal{M}$, the subspace of $\mathcal{R}^{n_k}$ that is accessible corresponds to $\mathcal{R}^{\vert m \vert}$ whose orthonormal basis vectors are given by $\boldsymbol{e}^k \;\; \forall \;\; k \in m$. Maximum velocity limits along each dimension impose further constraints on the set of control commands that are available in each mode. This constrained set of control commands available in mode $m$, denoted by $\mathcal{U}^m$, can be written as 
\begin{equation*}
\mathcal{U}^m = \{\boldsymbol{u} | \boldsymbol{u} \in \mathcal{R}^{\vert m \vert} ~ and ~ \norm{\boldsymbol{u}}_{\infty} \leq v_{lim} \} 
\end{equation*}
where $v_{lim}$ denotes the maximum velocity along any dimension $k$ and $\norm{\cdot}_{\infty}$ denotes the $L_\infty$ norm. 
%Therefore, the set $\mathcal{U}^m$ corresponds to a hypercube of dimensionality $\vert m \vert$ embedded in $\mathcal{R}^{n_k}$. In other words $\mathcal{U}^m$ is a $\vert m \vert$-cube with edge length equals $2c$. Without loss of generality, the velocity limits can be different for different control dimensions in which case $\mathcal{U}^m$ will correspond to a hypercuboid. Figure 1 represents this in a pictorial fashion for $\mathbb{R}^2$. 
% \begin{figure}[h]
%	\centering
%	\includegraphics[width= 0.8\hsize, height=0.2\vsize]{./figures/R2_ControlModes.eps}
%	\vspace{-0.35cm}
%	\caption{Illustration of 1D and 2D control modes in a two-dimensional control space, $\mathbb{R}^2$. Note that this space can be embedded in a higher dimensional space such as $\mathbb{R}^3$, $\mathbb{SE}(2)$ and $\mathbb{SE}(3)$. The 1D control modes are shown in blue and orange and the 2D control mode is shown as the green shaded area within the red square. The vertex set denotes the maximum magnitude control commands that can be issued when operating in a control mode. }
%	\label{fig:r2_modes}
%\end{figure}

% The disambiguation formalism developed in Section~\ref{ssec:disamb} is agnostic to the particular form of intent inference. However, the algorithm assumes that $\boldsymbol{p}(t)$ can be forward projected in time by iteratively applying the intent inference algorithm. 
\section{Intent Disambiguation}\label{sec:disamb}
This section describes how our autonomy counterfactually reasons about the intent disambiguation capabilities of different control modes by computing the information gain and content of the belief over goal utilizing a model-base approach in which the human is modeled is an adjustable rational agent that seeks to take the shortest distance path to the goal. 
We also describe the various types of intent inference schemes that we utilize in our experiements that work in conjunction with our disambiguation algorithm.

@@

In this subsection we show how the recursive Bayesian update of the probability distribution over goals is explicitly conditioned on the current active control mode and therefore can be leveraged to counterfactually reason about the disambiguation capabilities.

We treat the unknown goal $g$ as a random variable and  maintain a probability distribution $\boldsymbol{p}(t)$. We only consider a single evidence variable that corresponds to the human control command denoted as $\boldsymbol{u}_h$. The goal probability conditioned on the evidence variable can be written as 
\begin{equation}\label{eq:belief1}
b^t_g = p(g^t \;|\; \boldsymbol{u}_h^{0:t}) \propto p(\boldsymbol{u}_h^t\;|\;g^t,\boldsymbol{u}_h^{0:t-1})p(g^t|\boldsymbol{u}_h^{0:t-1})
\end{equation}
under the Markovian assumption that $(\boldsymbol{u}_h^t \perp \boldsymbol{u}_h^{0:t-1} | g^t)$, that is, at current timestep the human control command is independent of control commands at previous timesteps given the current goal). Marginalizing over $g^{t-1}$ Equation~\ref{eq:belief1} becomes
\begin{equation}\label{eq:belief2}
b^t_g = p(\boldsymbol{u}_h^t \;|\; g^t) \sum_{g^{t-1} \in \mathcal{G}}^{} p(g^t, g^{t-1} \; | \; \boldsymbol{u}_h^{0:t-1})
\end{equation}
and under the assumption that the conditional transition probability of changing to goal $g^t$ given that the goal is $g^{t-1}$ is independent of the control command history. Equation~\ref{eq:belief2}, therefore simplifies to
\begin{equation} \label{eq:recursive_belief}
\begin{split}
b^t_g & = p(\boldsymbol{u}_h^t \;|\; g^t) \sum_{g^{t-1} \in \mathcal{G}}^{} p(g^t | g^{t-1})  p(g^{t-1} | \boldsymbol{u}_h^{0:t-1})  \\
& = p(\boldsymbol{u}_h^t \;|\; g^t) \sum_{g^{t-1} \in \mathcal{G}}^{} p(g^t | g^{t-1}) b^{t-1}_{g}
\end{split}
\end{equation}
Marginalizing over $m$, the control mode variable, the human policy can be written as 
\begin{equation} \label{eq:marginal_mode}
p(\boldsymbol{u}_h^t \;|\; g^t) = \sum_{m \in \mathcal{M}}^{} p(\boldsymbol{u}_h^t \;|\; g^t, m)p(m)
\end{equation}
We can simplify Equation~\ref{eq:marginal_mode} using the critical piece of information that due to constrain of the control interface at any given time $t$ only one of the control modes in $\mathcal{M}$ can be active. That is, the probability distribution over modes at any given time $t$ reduces to a delta function. This can be written as
\begin{equation}\label{eq:mode_prob}
\begin{split}
p(m) & = 1 \;\;\;\; \text{if} \;\; m = m^t \\
& = 0 \;\;\;\; \text{otherwise}
\end{split}
\end{equation}
Using Equations~\ref{eq:marginal_mode} and \ref{eq:mode_prob} and under the assumption that user does not change the goal during task execution, Equation~\ref{eq:recursive_belief} can be further simplified as 
\begin{equation}\label{eq:final_belief}
	b^t_g = \underbrace{p(\boldsymbol{u}_h^t \;|\; g^t, m^t )}_{\substack{\text{policy conditioned} \\ \text{on mode}}}\;\underbrace{b^{t-1}_{g}}_{\substack{\text{belief at}\\ \text{$t-1$}}}
\end{equation}

@@

%\added{Rewrite this section in terms of math of Bayesian inference. And don't talk about need for intent disambiguation, but instead make a case for how inference can be affected by the control mode.Write basic equation. Under the constrain that user can only operate the robot in one mode at a time rewrite the equation conditioned on mode. Isolate the likelihood and explain the meaning of it. And how the fact that the likelihood differs for different modes can be exploited to come up with a mode selection paradigm. To characterize which one is good we use information gain and entropy measures. Check out Siddarths paper for recursive bayesian formulation. This is general to any recursive situation. Talk about human following a policy that noisily optimizes shortest distance in Cartesian space. Talk about counterfactual reasoning? That is, we are asking the question what if, the human operated in so and so mode, how would the probabilities update. }
%
%\added{Write the math in terms of mode conditioned recursive Bayesian update, under the constrain that the goal is not changing, and under the constrain that at any point in time only mode can be operative.
%}

Our aim is to develop a metric that will capture the ``disambiguation capability'' of a control dimension/mode. Subsequent user operation of the robot in the control mode with maximum disambiguation capability will likely help the system to perform better intent inference and may result in the autonomy providing more appropriate assistance.
At any given time $t$, the probability distributions over goals encodes the human's underlying intent. The time evolution of the probability distribution, however, depends on the choice of intent inference scheme and the various task relevant features and parameters that contribute to it.

>>>>

\begin{abstract}
The effectiveness of assistive robots is closely related to their ability to infer the user's needs and intentions \textit{unambiguously} and provide appropriate assistance quickly and accurately. In this paper, we proposed a mode selection paradigm that enhances the autonomy's intent inference capabilities by exploiting the fact that how the human operates the robot and reveals his/her intent is intrinsically conditioned on the control mode in operation. We formulate this as a problem of intent disambiguation in information theoretic terms.  We propose two different methods for enhancing intent inference via improved disambiguation using the information theoretic concepts of \textit{entropy} and \textit{KL divergence}. In our system, the autonomy maintain a probability distribution over goals (beliefs) and we characterize the disambiguation capabilities of the different control modes/dimensions utilizing a model-based approach to compute the average information gain and content. 
%The proposed algorithms characterize the disambiguation capabilities of the \added{different control modes?} system through the forward projection of probability distributions over goals, utilizing an intent inference scheme and a kinematics model, and computing the average information gain and content. 
Previous work has shown that the success of a disambiguation algorithm depends on a variety of factors and parameters. 
%To thoroughly investigate the impact of these various components, 
We present results from an extensive simulation-based study for both a point robot and physics-based simulation of a six degrees-of-freedom (DoF) robotic arm. Our results indicate that, compared to a baseline, the proposed disambiguation algorithms do enhance intent inference via faster intent disambiguation, which in turn allows autonomy assistance to step in earlier during task execution. We also find that goal inference is more accurate and the total amount of time assistance is engaged is higher.
\end{abstract}

@@

\section{Introduction}
Assistive machines such as robotic arms and smart wheelchairs have the potential to transform the lives of millions of people with motor impairments~\citep{laplante1992assistive}. These machines can promote independence, enhance the quality of lives, extend the mobility and manipulation capabilities of individuals and revolutionize the way people interact with society. 
%They can also help , thereby helping motor-impaired people perform activities of daily living in a more effective manner.

An assistive robotic machine is typically controlled using an interface such as a joystick, a switch-based head array or a sip-and-puff. These interfaces are low-dimensional, low-bandwidth and occasionally discrete. For this reason, at any given point in time during task execution they can only operate in a subset of the entire control space. These subsets are referred to as \textit{control modes}~\citep{simpson2008tooth}. To schedule and execute switches between control modes can be both mentally and physically demanding, and might be alleviated in part by the introduction of robotics autonomy. The efficacy of such autonomy-endowed machines relies on their ability to infer the users' needs and intentions, which is often a bottleneck for providing appropriate assistance accurately, confidently and quickly. Due to the dimensionality mismatch between high-dimensional robots and low-dimensional interfaces, the user is constrained to produce control commands that likely do not express their true underlying intent unambiguously. That is, the control interfaces act like filters that restrict the amount of information regarding the user's intent that is passed through to the autonomy. 
%Furthermore, sparsity and noise in the control commands make the inference process even harder prompting the need for robust intent inference formalisms.

In the context of assistive robotic manipulation, the first step of a task is often to \textit{reach} towards discrete objects in the environment. Therefore in this domain, \textit{intent inference} can be cast as a problem of maintaining and updating the probability distributions over all possible discrete goals (objects) in the workspace upon receiving fresh evidence from the human actions and environment. In Bayesian terminology, maintaining and updating the belief is the \textit{recursive Bayesian filtering} problem. Intent inference algorithms typically rely on various environmental cues and task-relevant features such as the robot and goal positions, human control commands and biometric measures~\citep{croft2003estimating} as their evidence variables. In contrast, due to reasons of user adoption and cost, our system matches the interfaces available on typical commercial assistive machines (such as powered wheelchairs) and so infers intent exclusively based on the information contained in the constrained control commands issued to the assistive machine.

Our system is inspired by the following insights:
\begin{itemize}
	\item At any given time, the user is constrained to operate the robot in a specific control mode. That is, the policy followed by the human while executing the reaching task is conditioned on the current active mode. 
	\item The posterior computation of the belief over goals depends on the human policy which indirectly depends on the active control mode. 
	\item Utilizing a model-based approach to forward project the beliefs over goals, the autonomy can \textit{counterfactually } reason about the information gain and content for different control modes. 
	\item Using this information theoretic characterization of control modes, the autonomy chooses the control mode with the highest intent disambiguation capability \textit{for} the human.
	\item Having the human provide control input within these information-rich control modes can likely improve the accuracy of intent inference. This, in turn, allows the autonomy to step in and \textit{assist the human achieve their desired goal earlier}. 
\end{itemize}
%Our insight is that, at any given time, there is a specific set of control modes that are more informative \textit{for} the autonomy than others. Therefore, by having the human provide control input within these information-rich modes, the possibility of inferring the human's intent accurately and unambiguously is increased. \added{Talk about this specifically in terms of how mode selection can improve intent inference. }

%Having the autonomy to step in earlier and provide accurate assistance is important, because when the intent inference mechanism infers the incorrect goal (or is simply too slow), the user and autonomy often end up competing instead of collaborating and can decrease user satisfaction and result in significantly worse performance. 
%than either the human or autonomy would produce on their own.

In this paper, we address the problem of accurate intent inference by characterizing control modes according to their information content and proposing a mode switch assistance paradigm that selects the control mode in which a user-initiated motion will \textit{maximally disambiguate} their intent. We formalize the problem of intent disambiguation within the framework of information theory. Specifically, we make use of the information theoretic notions of entropy and Kullback-Leibler (KL) divergence to characterize and quantify the information content---with respect to intent disambiguation---of each control mode.

@@

%Information theoretic approaches are widely utilized in the field of machine learning and robotics for optimal experiment design, for efficient data collection processes and for informing search strategies. 
Robot assistance schemes that seek to elicit more informative cues \textit{from} the human to clarify the underlying intent can be thought of as an information acquisition problem. Intent acquisition can leverage the underlying synergies and shared intentionality~\citep{tomasello2007shared} of human-robot teams and can be an active process in which the robot performs actions 
%(for example, selecting a control mode or executing a robot motion) 
that will nudge the human to reveal her/his intent more clearly~\cite{sadigh2016information}. Information theoretic metrics such as KL divergence can be utilized to identify regions of the sample space that will maximize information gain~\citep{tong2001active} and subsequently guide the data sampling process. 
%Sensing robots designed for automated exploration and data acquisition tasks can benefit from exploring more information-rich regions in the environment~\citep{atanasov2014information}. 
If the spatial distribution of information density is known \textit{a priori}, information maximization can be accomplished by maximizing the ergodicity of the robot's trajectory with respect to the underlying information density~\citep{miller2016ergodic}.

%For an assistive machine to provide appropriate kinds of assistance accurately and at the right time, it needs to have a good estimate of the human's underlying intent.
 Intent inference algorithms play a vital role in the success of an assistive system. Intent inference and recognition can be classified into two broad categories: heuristic approaches and model-based approaches. Heuristic approaches are often simpler and computationally light-weight and seek to find direct mappings between various task relevant features (such as motion cues) and the human's underlying intention~\citep{baker2007goal}. On the other hand, in model-based approaches the system maintains a model of how a human maps states to control actions. The model can either be learned from data or can be hand-designed based on domain knowledge. For example, the human can be modeled within the Partially Observable Markov Decision Process (POMDP)~\citep{taha2011pomdp} framework and is assumed to behave according to a control policy that maps states to actions. However, in model-based approaches incorporating the entire history of states requires estimation of joint probability distributions over past states which can become computationally expensive and intractable quickly.

When there is a mismatch between the dimensionality of the problem and the control interface users have to continuously shift their focus from the task at hand to the choice of control mode during task execution, thereby resulting in a higher cognitive load. Various mode switching assistance paradigms have been proposed to alleviate task effort. For example, an automatic time-optimal mode switch assistance has been proposed which has shown to significantly improve user satisfaction~\citep{herlant2016assistive}. In the area of myoelectric prosthetics, dynamic switching approaches that learn the most effective control option during task execution using temporal difference and reinforcement learning have also been proposed~\cite{pilarski2012dynamic}.

@@

% Section~\ref{ssec:notation} outlines the mathematical notation used in this paper. Section~\ref{ssec:set_modes} presents a set-theoretic treatment of control modes and Section~\ref{ssec:disamb} describes the disambiguation algorithm. The mathematical details of the intent inference paradigms are outlined in detail in Section~\ref{ssec:inference}.
\subsection{Probability Distribution Over Goals}\label{ssec:notation}
 In assistive robotic manipulation, intent inference is most commonly the process of estimating the user's intended goal out of a set of discrete objects in the environment~\citep{calli2015ycb}. The set of all candidate goals is denoted by $\mathcal{G}$ with $n_g = \vert\mathcal{G}\vert$.
% and let $g^i$ refer to the $i^{th}$ goal with $i \in [1,2,\dots, n_g]$. 
% At any time $t$, the likelihood of goal $g$ being the desired goal follows a categorical probability distribution whose sample space is $\mathcal{G}$. 
% That is, 
% \begin{equation*}
% g \sim \textbf{Cat}(n_g | \boldsymbol{p}(t)) = \prod_{i=1}^{n_g}p^{i}(t)^{[g = g^i]}
% \end{equation*}
%where, $[g = g^i]$ is the Iverson bracket and evaluates to 1 if $g = g^i$, 0 otherwise. 
At time $t$ the random variable $g^t \in \mathcal{G}$ denotes the user's intended goal and $b_{g^t}$ denotes the belief over the goals. 
% The probability $b_{g^t}$  also represents the robot's \textit{confidence} that goal $g$ is the human's intended goal at time $t$. 
%$\boldsymbol{p}(t)$ denotes the probability distribution over goals such that $\boldsymbol{p}(t) = [p_{g_1}^t, p_{g_2}^t,\dots, p_{g_{n_g}}^t)]^{T}$ where $p_{g_i}^t$ denotes the probability associated with goal $g_i$ at time $t$.

@@

 Let $\mathcal{M}$ denote the set of all control modes with $n_m = \vert\mathcal{M}\vert$. Additionally, let $m^i$ refer to the $i^{th}$ control mode where $i \in [1,2,\dots,n_m]$. Each control mode $m^i$ is a subset of $\mathcal{K}$ such that $\bigcup\limits_{i=1}^{n_m} m^i = \mathcal{K}$. The cardinality of each mode $m \in \mathcal{M}$, denoted by $\vert m \vert$, indicates the number of dimensions that can be controlled when operating in $m$.\footnote{Note that a dimension $k \in \mathcal{K}$ can be an element of multiple control modes and so it is possible that $m^i \cap m^j \neq \emptyset$.} Furthermore, the user can only operate in one of the $n_m$ control modes at any given time $t$. That is, for each $m \in \mathcal{M}$, the subspace of $\mathcal{R}^{n_k}$ that is accessible corresponds to $\mathcal{R}^{\vert m \vert}$ whose orthonormal basis vectors are given by $\boldsymbol{e}^k \;\; \forall \;\; k \in m$. Maximum velocity limits along each dimension impose further constraints on the set of control commands that are available in each mode. 
% This constrained set of control commands available in mode $m$, denoted by $\mathcal{U}^m$, can be written as 
%\begin{equation*}
%\mathcal{U}^m = \{\boldsymbol{u} | \boldsymbol{u} \in \mathcal{R}^{\vert m \vert} ~ and ~ \norm{\boldsymbol{u}}_{\infty} \leq v_{lim} \} 
%\end{equation*}
%where $v_{lim}$ denotes the maximum velocity along any dimension $k$ and $\norm{\cdot}_{\infty}$ denotes the $L_\infty$ norm. 
%Therefore, the set $\mathcal{U}^m$ corresponds to a hypercube of dimensionality $\vert m \vert$ embedded in $\mathcal{R}^{n_k}$. In other words $\mathcal{U}^m$ is a $\vert m \vert$-cube with edge length equals $2c$. Without loss of generality, the velocity limits can be different for different control dimensions in which case $\mathcal{U}^m$ will correspond to a hypercuboid. Figure 1 represents this in a pictorial fashion for $\mathbb{R}^2$. 
% \begin{figure}[h]
%	\centering
%	\includegraphics[width= 0.8\hsize, height=0.2\vsize]{./figures/R2_ControlModes.eps}
%	\vspace{-0.35cm}
%	\caption{Illustration of 1D and 2D control modes in a two-dimensional control space, $\mathbb{R}^2$. Note that this space can be embedded in a higher dimensional space such as $\mathbb{R}^3$, $\mathbb{SE}(2)$ and $\mathbb{SE}(3)$. The 1D control modes are shown in blue and orange and the 2D control mode is shown as the green shaded area within the red square. The vertex set denotes the maximum magnitude control commands that can be issued when operating in a control mode. }
%	\label{fig:r2_modes}
%\end{figure}

% The disambiguation formalism developed in Section~\ref{ssec:disamb} is agnostic to the particular form of intent inference. However, the algorithm assumes that $\boldsymbol{p}(t)$ can be forward projected in time by iteratively applying the intent inference algorithm. 
\section{Intent Disambiguation}\label{sec:disamb}
This section describes how our autonomy reasons about the intent disambiguation capabilities of different control modes by computing the information gain and content of the belief over goals. We utilize a model-based approach in which the human is modeled as an adjustable rational agent that seeks to take the shortest distance path to the goal. 
We also describe the various intent inference schemes that we utilize in our experiments, that work in conjunction with our disambiguation algorithm.

@@

In this subsection we show how the recursive Bayesian update of the probability distribution over goals is explicitly conditioned on the current active control mode and therefore can be leveraged to reason about disambiguation.
%the disambiguation capabilities.

%We treat the unknown goal $g$ as a random variable and  maintain a belief over the goals denoted as $b^t_g$. 
We only consider a single evidence variable that corresponds to the human control command denoted as $\boldsymbol{u}_h$. The goal probability conditioned on the evidence variable can be written as 
\begin{equation}\label{eq:belief1}
b_{g^t} = p(g^t \;|\; \boldsymbol{u}_h^{0:t}) \propto p(\boldsymbol{u}_h^t\;|\;g^t,\boldsymbol{u}_h^{0:t-1})p(g^t|\boldsymbol{u}_h^{0:t-1}) \;\;\;\;\;.
\end{equation}
Assuming that $(\boldsymbol{u}_h^t \perp \boldsymbol{u}_h^{0:t-1} | g^t)$ (Markovian assumption that the control command at the current timestep is independent of previous timesteps given the current goal)and marginalizing over $g^{t-1}$, Equation~\ref{eq:belief1} becomes
\begin{equation}\label{eq:belief2}
b_{g^t} = p(\boldsymbol{u}_h^t \;|\; g^t) \sum_{g^{t-1} \in \mathcal{G}}^{} p(g^t, g^{t-1} \; | \; \boldsymbol{u}_h^{0:t-1})
\end{equation}
We also assume that the conditional transition probability of changing to goal $g^t$ given that the goal is $g^{t-1}$ is independent of the control command history. Equation~\ref{eq:belief2}, therefore, simplifies to
\begin{equation}\label{eq:recursive_belief}
	b_{g^t} = p(\boldsymbol{u}_h^t \;|\; g^t) \sum_{g^{t-1} \in \mathcal{G}}^{} p(g^t | g^{t-1}) b_{g^{t-1}}
\end{equation}
%\begin{equation} \label{eq:recursive_belief}
%\begin{split}
%b^t_g & = p(\boldsymbol{u}_h^t \;|\; g^t) \sum_{g^{t-1} \in \mathcal{G}}^{} p(g^t | g^{t-1})  p(g^{t-1} | \boldsymbol{u}_h^{0:t-1})  \\
%& = p(\boldsymbol{u}_h^t \;|\; g^t) \sum_{g^{t-1} \in \mathcal{G}}^{} p(g^t | g^{t-1}) b^{t-1}_{g}
%\end{split}
%\end{equation}
Marginalizing over $m$, the control mode variable, the human policy $p(\boldsymbol{u}_h^t \;|\; g^t)$ can be written as 
\begin{equation} \label{eq:marginal_mode}
p(\boldsymbol{u}_h^t \;|\; g^t) = \sum_{m \in \mathcal{M}}^{} p(\boldsymbol{u}_h^t \;|\; g^t, m)p(m)
\end{equation}
We can simplify Equation~\ref{eq:marginal_mode} using the critical piece of information that, due to constraint of the control interface, at any given time $t$ only one of the control modes in $\mathcal{M}$ can be active. That is, the probability distribution over modes at any given time $t$ reduces to a delta function. This can be written as
\begin{equation}\label{eq:mode_prob}
	p(m) = \begin{cases}
			1 \;\;\;\; \text{if} \;\; m = m^t \\
			0 \;\;\;\; \text{otherwise}
			\end{cases}
\end{equation}
%\begin{equation}\label{eq:mode_prob}
%\begin{split}
%p(m) & = 1 \;\;\;\; \text{if} \;\; m = m^t \\
%& = 0 \;\;\;\; \text{otherwise}
%\end{split}
%\end{equation}
Using Equations~\ref{eq:marginal_mode} and \ref{eq:mode_prob} and under the assumption that user does not change the goal during task execution, Equation~\ref{eq:recursive_belief} can be further simplified as 
\begin{equation}\label{eq:final_belief}
	b_{g^t} = \underbrace{p(\boldsymbol{u}_h^t \;|\; g^t, m^t )}_{\substack{\text{human policy} \\ \text{conditioned on mode}}}\;\underbrace{b_{g^{t-1}}}_{\substack{\text{belief at}\\ \text{$t-1$}}}
\end{equation}

@@

\subsection{Algorithm: Information Theoretic Intent Disambiguation}

Our aim is to develop a metric that will capture the ``disambiguation capability'' of a control dimension/mode. 
From Equation~\ref{eq:final_belief} we can see that the time evolution of the probability distribution is sensitive to the user control command (denoted by $\boldsymbol{u_h}$), and that it will evolve differently as the user controls the robot in different control modes. By utilizing a model-based approach, the autonomy can reason about the time evolution of the belief $b_{g^t}$ for all control modes $m \in \mathcal{M}$ and then select the mode that offers the highest information gain and content. That is, motion within maximally disambiguating control dimensions serves as a mechanism to enhance the accuracy of intent inference.

The full algorithm is presented in Algorithm~\ref{alg:disamb}. We first perform a model-based projection of the probability distribution over goals (lines 5-9, Section~\ref{ssec:model_based}). We then compute the disambiguation capability $D_m$ of a control mode using information theoretic measures (lines 10-11) We present two different information theoretic measures for disambiguation: 1) information content in the projected probability distributions over goal as encoded by the \textit{entropy} of the distribution (Section~\ref{sssec:ent}) and 2) information gain as a result of the time evolution of the belief as encoded by the \textit{KL divergence} between the posterior and the prior distributions (Section~\ref{sssec:kl}).
